---
title: "HW 03"
author: "Nandakumar Kuthalaraja"
format:
  html:
    embed-resources: true
toc: true
---


```{r}
#| label: Setup
#| execute: 
#| warning: false
#| error: false
#| output: false



if (!require("pacman")) 
  install.packages("pacman")

pacman::p_load(tidyverse,gghighlight,openintro,lubridate,palmerpenguins,glue,scales,countdown,ggthemes,gt,openintro,ggrepel,patchwork,quantreg,janitor,colorspace,broom,fs,here,dsbox,ggridges,gtable,ggimage,ggpubr,cowplot,png,sysfonts,showtext,grid)

# set theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))

#Add Fonts as Needed
font_add("oldFont", "font/Wolfsmith.ttf")  
showtext_auto()

# set width of code output
options(width = 85)

# set figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 7,        # 7" width
  fig.asp = 0.618,      # the golden ratio
  fig.retina = 3,       # dpi multiplier for displaying HTML output on retina
  fig.align = "center", # center align figures
  dpi = 300             # higher dpi, sharper image
)
```


## 1 - Du Bois challenge.

```{r}
#| label: Du Bois challenge.
#| execute: 
#| warning: false

# Read Data 
income <- read_csv("data/income.csv")

# Recode final class label for line break
income <- income |>
  mutate(
    Class = recode(Class, "$1000 AND OVER" = "1000\nAND OVER")
  )

# left table
label_table <- income |>
  mutate(
    Class = factor(Class, levels = rev(c(
      "$100-200", "$200-300", "$300-400", "$400-500",
      "$500-750", "$750-1000", "1000\nAND OVER"
    ))),
    row = as.numeric(Class)
  ) |>
  ggplot(aes(y = Class)) +
  geom_tile(aes(x = 1.5, width = 2, height = 1), fill = NA, color = "black") +
  geom_segment(
    data = tibble(y = seq(0.5, 7.5, 1)),
    aes(x = 0.9, xend = 2.9, y = y, yend = y),
    inherit.aes = FALSE,
    color = "gray7", linewidth = 0.2
  ) +
  geom_segment(
    data = tibble(x = c(0.9, 1.95, 2.9)),
    aes(x = x, xend = x, y = 0.5, yend = 7.5),
    inherit.aes = FALSE,
    color = "gray7", linewidth = 0.2
  ) +
  geom_text(aes(x = 1, label = Class), hjust = 0, size = 3.2, family = "oldFont") +
  geom_text(aes(x = 2, label = paste0("$", Average_Income)), hjust = 0, size = 3.2, family = "oldFont") +
  scale_y_discrete(limits = rev(c(
    "$100-200", "$200-300", "$300-400", "$400-500",
    "$500-750", "$750-1000", "1000\nAND OVER"
  ))) +
  xlim(0.9, 2.9) +
  labs(title = NULL, x = NULL, y = NULL) +
  theme_void() +
  theme(
    plot.margin = margin(5, 0, 5, 5),
    axis.text.y = element_blank(),
    plot.background = element_rect(fill = NA, color = NA),  
    panel.background = element_rect(fill = NA, color = NA)
  )

data <- income |>
  pivot_longer(cols = Rent:Other, names_to = "type", values_to = "measurement") |>
  mutate(
    type = factor(type, levels = c("Rent", "Food", "Clothes", "Tax", "Other")),
    Class = factor(Class, levels = rev(c(
      "$100-200", "$200-300", "$300-400", "$400-500",
      "$500-750", "$750-1000", "1000\nAND OVER"
    )))
  ) 
data$measurement[data$measurement == 0] <- 0.1

category <- c(
  Rent = "#000000",
  Food = "#8663A3",
  Clothes = "#E89A8D",
  Tax = "slategray3",
  Other = "#D6C8B0"
)

bar_plot <- ggplot(data, aes(x = Class, y = measurement, fill = type)) +
  geom_bar(stat = "identity", width = 0.8) +
  geom_text(
    aes(label = ifelse(measurement >= 1, paste0(measurement, "%"), "")),
    position = position_stack(vjust = 0.5),
    size = 3,
    color = "white"
  ) +
  scale_y_reverse() +
  scale_fill_manual(values = category) +
  coord_flip() +
  labs(
    title = "", 
  x = NULL,
  y = NULL,
  caption = "FOR FURTHER STATISTICS RAISE THIS FRAME."
  ) +
 
  theme(
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.title = element_blank(),
    legend.position = "top",
    legend.direction = "horizontal",
    legend.justification = "center",
     plot.caption = element_text(
       family = "oldFont",
      hjust = 0.5,       
      size = 7,
      margin = margin(t = 10)
    ),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  
  )


combined_plot <- (label_table + bar_plot + plot_layout(widths = c(1.3, 3)))

final_plot <- ggbackground(combined_plot, "images/parchment_paper-1074131_1920.png")
print(final_plot)

```

## 2 - COVID survey - interpret

```{r}
#| label: COVID survey - interpret

```

This plot offers a comprehensive look at how medical and nursing students from various backgrounds responded to statements about the COVID-19 vaccine, using a Likert scale where lower scores signify greater agreement with positive vaccine sentiments

### Asian students expressed more concern about safety/side effects of Vaccine

Seems Asian students have a higher mean score than most other groups. This reflects the dynamic Cultural aspects & maybe historical experiences with medical procedures.

Think it is little disappointing compared to, how they had high compliance with public health measures otherwise.

### Black/African students report higher agreement recommending vaccine

Despite broader societal data showing lower vaccine confidence among Black Americans (due to historic medical mistreatment and systemic inequities), this subgroup shows Strong agreement with recommedning vaccine.

Only reason I believe this is as thy are healthcare students, who may have more access to scientific information.

### Medical students tend to be more confident than nursing students

Medical students show slightly lower (more confident) scores in: The scientific vetting process & on trusting the information received

Medical students may have had more exposure to the bio-medical research or vaccine development process during training, resulting in higher trust in science. Agree with this assessment.

## 3 - COVID survey - reconstruct

```{r}
#| label: COVID survey - reconstruct
#| execute: 
#| warning: false
#| error: false

covid <- read_csv("data/covid-survey.csv")
#dim(covid)
### Old Dimension
#Re-Aligning to row1
covid <- read_csv("data/covid-survey.csv", skip=1)
dim(covid)

```
```{r}


cols_to_check <- setdiff(names(covid), "response_id")
covid <- covid |>
  filter(!if_all(all_of(cols_to_check), is.na))

dim(covid)

```


```{r}
vax_levels <- c("0" = "No", "1" = "Yes")
profession_levels <- c("0" = "Medical", "1" = "Nursing")
gender_levels <- c("0" = "Male", "1" = "Female", "3" = "Non-binary third gender", "4" = "Prefer not to say")
race_levels <- c(
  "1" = "American Indian/Alaskan Native", 
  "2" = "Asian", 
  "3" = "Black/African American", 
  "4" = "Native Hawaiian/Other Pacific Islander", 
  "5" = "White"
)
ethnicity_levels <- c("1" = "Hispanic/Latino", "2" = "Non-Hispanic/Non-Latino")
age_bin_levels <- c("0" = "<20", "20" = "21-25", "25" = "26-30", "30" = ">30")

covid <- covid |>
  mutate(
    exp_already_vax = recode(exp_already_vax, !!!vax_levels),
    exp_flu_vax = recode(exp_flu_vax, !!!vax_levels),
    exp_profession = recode(exp_profession, !!!profession_levels),
    exp_gender = recode(exp_gender, !!!gender_levels),
    exp_race = recode(exp_race, !!!race_levels),
    exp_ethnicity = recode(exp_ethnicity, !!!ethnicity_levels),
    exp_age_bin = recode(exp_age_bin, !!!age_bin_levels)
  )
dim(covid)

```


```{r}
covid_survey_longer <- covid |>
  pivot_longer(
    cols = starts_with("exp_"),
    names_to = "explanatory",
    values_to = "explanatory_value"
  ) |>
  filter(!is.na(explanatory_value)) |>
  pivot_longer(
    cols = starts_with("resp_"),
    names_to = "response",
    values_to = "response_value"
  )

covid_survey_longer
```


```{r}
covid_survey_longer <- covid_survey_longer |>
  mutate(response_value = as.numeric(response_value))

sapply(covid_survey_longer, class)

```


```{r}

covid_survey_summary <- covid_survey_longer |>
  group_by(explanatory, explanatory_value, response) |>
  summarise(mean = mean(response_value, na.rm = TRUE),
            low = quantile(response_value, 0.1, na.rm = TRUE),
            high = quantile(response_value, 0.9, na.rm = TRUE))

covid_survey_summary

```

```{r}

covid_survey_summary_all <- covid_survey_longer |>
  group_by(response) |>
  summarise(mean = mean(response_value, na.rm = TRUE),
            low = quantile(response_value, 0.1, na.rm = TRUE),
            high = quantile(response_value, 0.9, na.rm = TRUE)
            ) |>
  mutate(
    explanatory = "All",
    explanatory_value = ""
  )

covid_survey_summary_all$explanatory_value <- as.factor(as.character(covid_survey_summary_all$explanatory_value))

sapply(covid_survey_summary_all, class)

covid_survey_summary_all
```


```{r}
covid_survey_summary_stats <- bind_rows(covid_survey_summary_all, covid_survey_summary)

covid_survey_summary_stats
```

```{r}

covid_survey_summary_stats <- covid_survey_summary_stats |>
  mutate(
    explanatory_value = fct_relevel(explanatory_value, ">30", "26-30", "21-25", "<20"),
    explanatory_value = fct_relevel(explanatory_value, "Female", "Male", "Non-binary third gender", "Prefer not to say"),
    explanatory_value = fct_relevel(explanatory_value, "American Indian/Alaskan Native", "Asian", "Black/African American", "Native Hawaiian/Other Pacific Islander", "White"),
    explanatory_value = fct_relevel(explanatory_value, "Non-Hispanic/Non-Latino", "Hispanic/Latino"),
    explanatory_value = fct_relevel(explanatory_value, "Medical", "Nursing"),
    explanatory_value = fct_relevel(explanatory_value, "No", "Yes")
  )


covid_survey_summary_stats <- covid_survey_summary_stats |>
  mutate(
    explanatory = fct_relevel(explanatory,
      "All",
      "exp_age_bin", "exp_gender", "exp_race", "exp_ethnicity",
      "exp_profession", "exp_already_vax", "exp_flu_vax"
    )
  )

covid_survey_summary_stats <- covid_survey_summary_stats |>
  mutate(
    explanatory = recode(explanatory,
      "exp_age_bin" = "Age",
      "exp_gender" = "Gender",
      "exp_race" = "Race",
      "exp_ethnicity" = "Ethnicity",
      "exp_profession" = "Profession",
      "exp_already_vax" = "Had COVID vaccine",
      "exp_flu_vax" = "Had flu vaccine this year"
    ),
    response = fct_relevel(response,
      "resp_safety",
      "resp_feel_safe_at_work",
      "resp_concern_safety",
      "resp_confidence_science",
      "resp_trust_info",
      "resp_will_recommend"
    ),
    response = recode(response,
      "resp_safety" = "Based on my understanding, I believe the vaccine is safe",
      "resp_feel_safe_at_work" = "Getting the vaccine will make me feel safer at work",
      "resp_concern_safety" = "I am concerned about the safety and side effects of the vaccine",
      "resp_confidence_science" = "II am confident in the scientific vetting process for the new COVID vaccines",
      "resp_trust_info" = "I trust the information that I have received about the vaccines",
      "resp_will_recommend" = "I will recommend the vaccine to family, friends, and community members"
    )
  )

covid_survey_summary_stats
```

```{r}
ggplot(covid_survey_summary_stats, aes(x = mean, y = explanatory_value)) +
  geom_point(size = 0.80) +
  geom_errorbarh(aes(xmin = low, xmax = high), height = 0.3) +
  facet_grid(
    rows = vars(explanatory),
    cols = vars(response),
    scales = "free_y",
    space = "free_y",
    labeller = labeller(
      explanatory = label_wrap_gen(15),
      response = label_wrap_gen(15)
    )
  ) +
  scale_x_continuous(breaks = 1:5) +
  labs(
    x = "Mean Likert score\n(Error bars show 10th–90th percentile)",
    y = NULL
  ) +
  theme(
    strip.text = element_text(size = 6),
    strip.text.y = element_text(angle = 0),
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),
    axis.title.x = element_text(size = 8),
    panel.spacing = unit(0.2, "lines"),
    panel.grid = element_blank(),
    strip.background = element_rect(fill = "gray90", color = "black")
  )
```

## 4 - COVID survey - re-reconstruct

```{r}
#| execute: 
#| warning: false
#| error: false
covid_survey_summary_new <- covid_survey_longer |>
  group_by(explanatory, explanatory_value, response) |>
  summarise(mean = mean(response_value, na.rm = TRUE),
            low = quantile(response_value, 0.25, na.rm = TRUE),
            high = quantile(response_value, 0.75, na.rm = TRUE))
```

```{r}
#| execute: 
#| warning: false
#| error: false



covid_survey_summary_all_new <- covid_survey_longer |>
  group_by(response) |>
  summarise(mean = mean(response_value, na.rm = TRUE),
            low = quantile(response_value, 0.25, na.rm = TRUE),
            high = quantile(response_value, 0.75, na.rm = TRUE)
            ) |>
  mutate(
    explanatory = "All",
    explanatory_value = ""
  )

covid_survey_summary_all_new$explanatory_value <- as.factor(as.character(covid_survey_summary_all_new$explanatory_value))

sapply(covid_survey_summary_all_new, class)

```

```{r}
#| execute: 
#| warning: false
#| error: false
covid_survey_summary_stats_new <- bind_rows(covid_survey_summary_all_new, covid_survey_summary_new)

covid_survey_summary_stats_new
```

```{r}
#| execute: 
#| warning: false
#| error: false
covid_survey_summary_stats_new <- covid_survey_summary_stats_new |>
  mutate(
    explanatory_value = fct_relevel(explanatory_value, ">30", "26-30", "21-25", "<20"),
    explanatory_value = fct_relevel(explanatory_value, "Female", "Male", "Non-binary third gender", "Prefer not to say"),
    explanatory_value = fct_relevel(explanatory_value, "American Indian/Alaskan Native", "Asian", "Black/African American", "Native Hawaiian/Other Pacific Islander", "White"),
    explanatory_value = fct_relevel(explanatory_value, "Non-Hispanic/Non-Latino", "Hispanic/Latino"),
    explanatory_value = fct_relevel(explanatory_value, "Medical", "Nursing"),
    explanatory_value = fct_relevel(explanatory_value, "No", "Yes")
  )


covid_survey_summary_stats_new <- covid_survey_summary_stats_new |>
  mutate(
    explanatory = fct_relevel(explanatory,
      "All",
      "exp_age_bin", "exp_gender", "exp_race", "exp_ethnicity",
      "exp_profession", "exp_already_vax", "exp_flu_vax"
    )
  )

covid_survey_summary_stats_new <- covid_survey_summary_stats_new |>
  mutate(
    explanatory = recode(explanatory,
      "exp_age_bin" = "Age",
      "exp_gender" = "Gender",
      "exp_race" = "Race",
      "exp_ethnicity" = "Ethnicity",
      "exp_profession" = "Profession",
      "exp_already_vax" = "Had COVID vaccine",
      "exp_flu_vax" = "Had flu vaccine this year"
    ),
    response = fct_relevel(response,
      "resp_safety",
      "resp_feel_safe_at_work",
      "resp_concern_safety",
      "resp_confidence_science",
      "resp_trust_info",
      "resp_will_recommend"
    ),
    response = recode(response,
      "resp_safety" = "Based on my understanding, I believe the vaccine is safe",
      "resp_feel_safe_at_work" = "Getting the vaccine will make me feel safer at work",
      "resp_concern_safety" = "I am concerned about the safety and side effects of the vaccine",
      "resp_confidence_science" = "II am confident in the scientific vetting process for the new COVID vaccines",
      "resp_trust_info" = "I trust the information that I have received about the vaccines",
      "resp_will_recommend" = "I will recommend the vaccine to family, friends, and community members"
    )
  )

covid_survey_summary_stats_new
```

```{r}
#| execute: 
#| warning: false
#| error: false
ggplot(covid_survey_summary_stats_new, aes(x = mean, y = explanatory_value)) +
  geom_point(size = 0.80) +
  geom_errorbarh(aes(xmin = low, xmax = high), height = 0.3) +
  facet_grid(
    rows = vars(explanatory),
    cols = vars(response),
    scales = "free_y",
    space = "free_y",
    labeller = labeller(
      explanatory = label_wrap_gen(15),
      response = label_wrap_gen(15)
    )
  ) +
  scale_x_continuous(breaks = 1:5) +
  labs(
    x = "Mean Likert score\n(Error bars show 25th–75th percentile)",
    y = NULL
  ) +
  theme(
    strip.text = element_text(size = 6),
    strip.text.y = element_text(angle = 0),
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),
    axis.title.x = element_text(size = 8),
    panel.spacing = unit(0.2, "lines"),
    panel.grid = element_blank(),
    strip.background = element_rect(fill = "gray90", color = "black")
  )
```

This Plot differs majorly from the previous plot.

The change of narrowing the error bar range between 25th-75th percentile, have significantly changed the plot as error bars that previously spanned nearly the full Likert scale are now more tightly centered around their respective means.

The new plot reveals more about the distribution of responses with reduced error limit.

## 5 - COVID survey - another view


```{r}
#| execute: 
#| warning: false
#| error: false

covid_data_subset = subset(covid, select = -c(exp_profession, exp_flu_vax, exp_gender, exp_race, exp_ethnicity, exp_age_bin))

covid_data_subset <- covid_data_subset |>
    pivot_longer(
    cols = starts_with("resp_"),
    names_to = "response",
    values_to = "response_value"
    ) |>
  filter(!is.na(exp_already_vax))

sapply(covid_data_subset, class)

covid_data_subset$response_value <- as.numeric(as.character(covid_data_subset$response_value))

covid_data_subset_summary <- covid_data_subset |>
  group_by(response, exp_already_vax) |>
  summarize(mean = mean(response_value, na.rm = TRUE))

covid_data_subset_summary <- covid_data_subset_summary |>
  mutate(mean_adj = ifelse(exp_already_vax == "No", -mean, mean))

covid_data_subset_summary <- covid_data_subset_summary |>
  mutate(
          response = recode(response,
      "resp_safety" = "Based on my understanding, I believe the vaccine is safe",
      "resp_feel_safe_at_work" = "Getting the vaccine will make me feel safer at work",
      "resp_concern_safety" = "I am concerned about the safety and side effects of the vaccine",
      "resp_confidence_science" = "II am confident in the scientific vetting process for the new COVID vaccines",
      "resp_trust_info" = "I trust the information that I have received about the vaccines",
      "resp_will_recommend" = "I will recommend the vaccine to family, friends, and community members"
    ))
covid_data_subset_summary$response_wrap = str_wrap(covid_data_subset_summary$response, width = 45)


covid_data_subset_summary |>
  ggplot(aes(x = mean_adj, y = reorder(response_wrap, -mean_adj), fill = exp_already_vax)) +
  geom_col(position = "identity", width = 0.6) +
  labs(
    y = "Responses", 
    x = "Mean",
    title = "Diverging Bar Chart",
    subtitle = "1 (strongly agree) to 5 (strongly disagree)"
  ) +
  scale_x_continuous(
    labels = abs,
    limits = c(-5, 5),
    breaks = c(-5, -4, -3, -2, -1, 0, 1, 2, 3, 4, 5)
    ) +
  annotate(
    geom = "text",
    x = 0,
    y = Inf,
    label = "Not already vaccinated  ",
    size = 3,
    color = "darkgreen",
    hjust = 1
  ) +
   annotate(
    geom = "text",
    x = 0,
    y = Inf,
    label = "  Already vaccinated ",
    size = 3,
    color = "lightblue",
    hjust = 0
  ) +
  scale_fill_manual(values=c("darkgreen","lightblue")) +
  theme(
    panel.grid.minor.x = element_blank(),
    legend.position = "none",
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 7),
    plot.title.position = "plot",
    plot.subtitle = element_text(size = 10)
  ) +
  coord_cartesian(clip = "off")

```

```{r}


```



```{r}


```
